---
title: "TESA Workshop on ageing"
author: "Chairs: Daniel Ricard (GFC) and Peter Comeau (BIO)"
date: "2023-01-31 to 2023-02-02"
output: beamer_presentation
urlcolor: blue
header-includes:
- \AtBeginDocument{\title[TESA Workshop on age determination]{TESA workshop - Best practices in age determination}}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(gulf))
```

## What is TESA?

[TESA webpage on GCpedia](https://www.gcpedia.gc.ca/wiki/DFO_Science_TESA_ETES)


## Motivations and steps leading to this workshop

- initial idea proposed to TESA committee back in 2021
- age-based assessments are commonly used in providing science advice
- many new DFO scientists might not be familiar with best practices in fish age determination
- catch-at-age information is an important component of age-based population models

- there is no community of DFO scientists with age determination tasks beyond regional groups


## Land acknowledgment

The city of Moncton is located in traditional Mi'kmaq territory


[Elsipogtog Mi'kmaq cultural centre](https://heritagepathtour.com/)

[Site d’interprétation Micmac de Gespeg](http://www.micmacgespeg.ca/)


## Workshop participants

- potential participants initially identified regionally through TESA
- participants that received and accepted the Outlook meeting request
- snowball effect of having a virtual event, and the fact that we didn't turn anyone down
- please make sure that your presence is correctly noted in [GitHub README page](https://github.com/TESA-workshops/ageing-best-practices-2023)

- __please make sure that your presence is correctly documented in the draft workshop proceedings__


```{r participants, include=FALSE}
par.df <- readxl::read_xlsx("../planning/workshop-participants.xlsx", sheet = "for-README", skip=1)
par.df <- par.df[2:nrow(par.df),c(2,4,6,8)]

```
```{r, results='asis'}
knitr::kable(aggregate(Number~Region,par.df,length))
```

## External experts
\includegraphics[width=0.60\textwidth]{julie.png}

\includegraphics[width=1.0\textwidth]{lisa.png}

## External expert that says hello

\begin{columns}
\column{0.3\textwidth}
\begin{small}
Hi Dan and Peter,

As you can see from these photos, I am having a miserable time here, so the least you can do is to toast me in my time of despair...

All the best,

Steve
\end{small}
\column{0.7\textwidth}
\begin{center}
\includegraphics[width=0.8\textwidth]{campana.png}
\includegraphics[width=0.5\textwidth]{campana2.png}
\end{center}

\end{columns}

## Training vs. workshop

- Share your ideas and experiences
- Ask questions that may have an answer within the participants' collective knowledge 
- Formulating best practices can also mean identifying bad practices
- Our shared goal is to obtain age estimates that are unbiased and precise

- Collective experience of workshop participants is vast


## Virtual meeting

- In person participants
  - Rotating rapporteur role
  - Use your microphone when talking
  
- Guidelines for MS Teams 
  - Raise hand to make an intervention
  - Turn on your camera when speaking (unless you have bandwidth issues)
  - Unmute yourself to talk

- We will pause and repeat if there are connection issues with the external experts


## Day 1 - Tuesday January 31 2023

- Basics of age determination by Peter Comeau

- [Obtaining and cataloguing otoliths, preparation, storage and inventory](./preparation-storage-inventory.pdf)

- Plenary talk by Tracey Loewen (DFO Central and Arctic)

- Plenary talk by Stephen Wischniowski (DFO Pacific)

- [Validation, verification and calibration](./validation-verification-calibration.pdf)

- Age determination in a production ageing environment
  - Kim Emond and Hélène Dionne (IML) - Age determination of Atlantic Herring in the Québec region
  - Sylvie Robichaud and Karen Robertson (GFC) - Age determination of Atlantic Herring in the Gulf region
  - Tania Davignon-Burton (BIO) - Reconciling dreams, expectations and reality in a production ageing environment

- Towards establishing a sclerochronology community within DFO

## Roundtable of introductions for in person participants

- Name, position and region, mandated tasks that are making you attend this workshop


## Community building ideas

- Community of Age Reading Experts ([CARE](https://care.psmfc.org/)) as a test case
- Identify the great value of otolith exchanges between labs
- Develop a shared understanding of what physical and digital samples we have, and what they were/are/will be used for
- Shared References related to age determination
- ...

## Day 2 - Wednesday February 01 2023

- Practical sessions / tour of the GFC ageing lab

- SmartDots: a tool for annotations of otolith images, supporting otolith exchanges between labs, and quality-control of ages (Julie Davies)

- Taking good images of otoliths (Karen Robertson and others)

- Archiving of digital images and integration with SmartDots (David Fishman and Daniel Ricard)

- SmartDots reporting features - example for American Plaice reference collection (Daniel Ricard and Julie Davies)

- [Growth increment measurements from annotated otolith images (Daniel Ricard)](images-SmartDots-increments..pdf)


## Day 3 - Thursday February 02 2023

- Plenary talk by Lisa Ailloud (NOAA NMFS SEFSC)

- Plenary talk by Andrea Perreault (NAFC) 

- [Growth models](./growth-models.pdf)

- [Age-length keys and catch-at-age matrices](./age-length-keys.pdf)

- CatchR: Methods developed to compute age-length keys for the northern Gulf of St. Lawrence Atlantic Cod, and prospects for other age-based assessments

- Kim Emond and Hélène Dionne (IML) - Catch-at-age of commercial herring landings and numbers-at-age from acoustics surveys

- [Past, present and future of age determination](./past-present-future.pdf)

## Recommendations


## Proceedings draft document

- [draft DFO Technical Report](../proceedings/_book/techreport-english.pdf)


## Homework results - by region

```{r homework0, include=FALSE}
in.df <- readxl::read_xlsx("../planning/Ageing-workshop-regional-homework-all-regions.xlsx", skip=3)
in.df$ageing.structure <- in.df$`Ageing structure`
```

```{r homework1}
library(ggplot2)
g <- ggplot(data=in.df, aes(factor(Region))) +
  geom_bar()
g
```

## Homework results - by ageing structure

```{r homework2}
g <- ggplot(data=in.df, aes(factor(ageing.structure))) +
  geom_bar()
g
```
