---
title: "TESA Workshop on ageing"
author: "Chairs: Daniel Ricard (GFC) and Peter Comeau (BIO)"
date: "2023-01-31 to 2023-02-02"
output: beamer_presentation
urlcolor: blue
header-includes:
- \AtBeginDocument{\title[TESA Workshop on ageing]{TESA workshop - Best practices in ageing}}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(gulf))
```

## Day 1 - Tuesday January 31 2023

- Basics of ageing by Peter Comeau

- [Obtaining and cataloguing otoliths, preparation, storage and inventory](./preparation-storage-inventory.pdf)

- Plenary talk by Tracey Loewen (DFO Central and Arctic)

- Plenary talk by Stephen Wischniowski (DFO Pacific)

- [Validation, verification and calibration](./validation-verification-calibration.pdf)

- Age determination in a production ageing environment
  - Kim Emond and Hélène Dionne (IML) - Age determination of Atlantic Herring in the Québec region
  - Sylvie Robichaud and Karen Robertson (GFC) - Age determination of Atlantic Herring in the Gulf region
  - Tania Davignon-Burton (BIO) - Reconciling dreams, expectations and reality in a production ageing environment

- Towards establishing a sclerochronology community within DFO

## Community building ideas

- Community of Age Reading Experts (CARE) as a test case
- Identify the great value of otolith exchanges between labs
- Develop a shared understanding of what physical and digital samples we have, and what they were/are/will be used for

## Day 2 - Wednesday February 01 2023

- Practical sessions / tour of the GFC ageing lab

- SmartDots: a tool for annotations of otolith images, supporting otolith exchanges between labs, and quality-control of ages (Julie Davies)

- Taking good images of otoliths (Karen Robertson and others)

- Archiving of digital images and integration with SmartDots (David Fishman and Daniel Ricard)

- SmartDots reporting features - example for American Plaice reference collection (Daniel Ricard and Julie Davies)

- [Growth increment measurements from annotated otolith images (Daniel Ricard)](images-SmartDots-increments..pdf)


## Day 3 - Thursday February 02 2023

- Plenary talk by Lisa Ailloud (NOAA NMFS SEFSC)

- Plenary talk by Andrea Perreault (NAFC) 

- [Growth models](./growth-models.pdf)

- [Age-length keys and catch-at-age matrices](./age-length-keys.pdf)

- CatchR: Methods developed to compute age-length keys for the northern Gulf of St. Lawrence Atlantic Cod, and prospects for other age-based assessments

- Kim Emond and Hélène Dionne (IML) - Catch-at-age of commercial herring landings and numbers-at-age from acoustics surveys

- [Past, present and future of age determination](./past-present-future.pdf)

## Recommendations


## Proceedings draft document

- [draft DFO Technical Report](../proceedings/_book/techreport-english.pdf)


## Homework results - by region

```{r homework0, include=FALSE}
in.df <- readxl::read_xlsx("../planning/Ageing-workshop-regional-homework-all-regions.xlsx", skip=3)
in.df$ageing.structure <- in.df$`Ageing structure`
```

```{r homework1}
library(ggplot2)
g <- ggplot(data=in.df, aes(factor(Region))) +
  geom_bar()
g
```

## Homework results - by ageing structure

```{r homework2}
g <- ggplot(data=in.df, aes(factor(ageing.structure))) +
  geom_bar()
g
```
